---
title: "Profiling Spatial Analysis"
author: "Job Spijker"
output: 
  pdf_document:
    highlight: tango
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: yes
urlcolor: blue    
---

```{r}

remotes::install_github("gearslaboratory/gdalUtils")
remotes::install_github("jspijker/datafile")

# init
library(rgdal)
library(gdalUtils)
library(sf)
library(microbenchmark)
library(raster)
library(datafile)
datafileInit()

```



# Introduction

We first define a bunch of benchmarks. These are a collection of
functions with very basic spatial I/O operations. Then we run all the
benchmarks together

# Raster analysis
 Speed test using raster package
 
## create raster
First simple test, just the creation of a large raster

```{r}


# Create raster, use Dutch RD extent
createRast <- function(xmin=0,xmax=280000,
                        ymin=300000,ymax=625000,
                        resolution=25) {

    xcells=(xmax-xmin)/resolution
    ycells=(ymax-ymin)/resolution

    x <- raster::raster(xmn=xmin,
                xmx=xmax,
                ymn=ymin,
                ymx=ymax,
                nrows=ycells,
                ncols=xcells,
                crs=CRS("+init=epsg:28992")
                )


    #x[] <- 1:ncell(x)
    x[] <- round(runif(n = ncell(x), min = 0, max = 1000))
    return(x)
}


```

## create raster, write to disk and read 

```{r}

write_rast <- function() {
    x1 <- createRast()
    fname <- datafile("rast")
    writeRaster(x1, fname, overwrite = TRUE)
    x2 <- raster(fname)

}


```

# create rasters, stack and stats

Create a few rasters, stack them and calculate some summary
statistics. We use smaller rasters because of memory limitations

```{r}
rasts <- paste0("rast", 1:9)
for(i in rasts) {
    x <- createRast(xmax = 100000, ymin = 30000, ymax = 130000)
    fname <- datafile(i)
    writeRaster(x, fname, overwrite = TRUE)
    rm(x)
}

rast_lst <- list()
for(i in rasts) {
    rast_lst[[i]] <- raster(datafile(i))
}

calc_stats <- function(l = rast_lst) {
    s <- stack(l)
    summary(s[])
    rm(s)
}




```

# vectors

```{r}

#wfsuri <- "WFS:https://geodata.nationaalgeoregister.nl/wijkenbuurten2020/wfs?request=GetCapabilities&service=wfs"

wfsuri <- "WFS:https://service.pdok.nl/cbs/wijkenbuurten/2020/wfs/v1_0?request=getcapabilities&service=wfs"

# get list of existing layers in WFS
layers <- ogrListLayers(wfsuri)
buurten <- layers[grep("cbs_buurten", layers)]
buurten_layername <- gsub(".*:", "", buurten, perl=TRUE)
fname_layer <- datafile(paste0(buurten_layername, ".gpkg"))
fname_tmplayer <- datafile(paste0("tmp_", buurten_layername, ".gpkg"))

# Only download layer from NGR if file does not exists
if(!file.exists(fname_layer)) {
    cat("Dowloading", buurten_layername, "\n")
    ogr2ogr(wfsuri, fname_layer,
            layer = buurten,
            f = "GPKG")

    ogr2ogr(fname_layer, fname_tmplayer, buurten, nlt = "MULTIPOLYGON")
}

read_write_sf <- function(fname = fname_layer) {

    fname_tmp <- datafile("tmp.gpkg")
    x <- st_read(fname_layer)
    st_write(x, fname_tmp, quit = TRUE)
    unlink(fname_tmp)
}



```

# Run Benchmarks

```{r}

b <- microbenchmark(createRast(),
                    write_rast(),
                    calc_stats(),
                    read_write_sf()
)
b


```


#cleanup

```{r}
rm(x); gc()
removeTmpFiles(h=0)
```

